---
title: "Bootstrap Omnibus Test for Project G"
author: "Linda Wei"
output: html_notebook
---

#### Bootstrap Omnibus Test inference 

A bootstrap function is conducted to generate 10000 resamples for each ESE group with replacements from the original data. We will compute the statistics of interest on each resample. These statistics are then used to estimate the sampling distribution of the statistic of interest and to construct confidence intervals.

```{r echo=FALSE, message=FALSE}
library(here)
school_data <- read.csv(here("data", "hsb2.csv"))
source(here("function", "bootstrap_means.R"))

# separate the data by ses
ses1 <- subset(school_data, ses == 1)$write
ses2 <- subset(school_data, ses == 2)$write
ses3 <- subset(school_data, ses == 3)$write

# bootstrap the means for each group
means1 <- bootstrap_means(ses1)
means2 <- bootstrap_means(ses2)
means3 <- bootstrap_means(ses3)

# calculate the confidence intervals
ci1 <- c(sort(means1)[251], sort(means1)[9750])
ci2 <- c(sort(means2)[251], sort(means2)[9750])
ci3 <- c(sort(means3)[251], sort(means3)[9750])

ci_table <- data.frame(SES = c("SES 1", "SES 2", "SES 3"),
                       Lower_CI = c(ci1[1], ci2[1], ci3[1]),
                       Upper_CI = c(ci1[2], ci2[2], ci3[2]))
print(ci_table)


# conduct the hypothesis test
alpha <- 0.05
ses_anova <- aov(school_data$write ~ school_data$ses, data = school_data)
f_stat <- summary(ses_anova)[[1]][["F value"]][1]
p_val <- summary(ses_anova)[[1]][["Pr(>F)"]][1]
ses_mean <- aggregate(school_data$write ~ school_data$ses, data = school_data, FUN = mean)
print(ses_mean)
```
The mean writing score for the "low" SES group is estimated to be between 47.89 and 53.32, with 95% confidence.
The mean writing score for the "middle" SES group is estimated to be between 50.06 and 53.75, with 95% confidence.
The mean writing score for the "high" SES group is estimated to be between 53.45 and 58.19, with 95% confidence.

While there is some overlap between the confidence intervals for the "low" and "middle" SES groups, there is no overlap between the confidence intervals for any two SES groups at the 95% level. This suggests that there are statistically significant differences in mean writing scores between all three groups, with the "low" SES group having the lowest estimated mean writing score, followed by the "middle" SES group, and then the "high" SES group. However, we should note that these are only estimates based on the observed data, and there is still some uncertainty associated with these estimates due to sampling variability.

#### Tukey pairwise confidence interval

Based on the bootstrap test result, the overall CI does not contain zero, it indicates that there is a statistically significant difference between the means of the three ses groups. However, it does not provide information on which specific pairs of means differ significantly from each other.Therefore, a post-hoc test Tukey pairwise comparison (Tukey HSD test) is conducted to determine which specific pairs of means differ significantly from each other, while controlling for the overall type I error rate.

First, a Tukey pairwise test is conducted based on the hsb2.csv data.
```{r echo=FALSE, message=FALSE}
# prepare data for ANOVA
write_data <- school_data[, c(4, 8)]

write_data$ses <- factor(write_data$ses,
                          levels = c(1, 2, 3))

# ANOVA
model <- aov(write_data$write ~ write_data$ses, data = write_data)
print(summary.aov(model))

# TukeyHSD
tukey_data <- TukeyHSD(model)

print(tukey_data)
```
The ANOVA table shows that there is a statistically significant difference in the mean writing scores of the groups defined by the variable ses (which stands for socioeconomic status) at the alpha level of 0.05. The F-value is 4.97 and the p-value is 0.00784.

The Tukey multiple comparisons of means table shows the results of pairwise comparisons between the three groups (SES levels 1, 2, and 3). Based on the table, we can conclude that there is a statistically significant difference between the mean writing scores of students with SES levels 1 and 3 (p adj = 0.0114079) and between students with SES levels 2 and 3 (p adj = 0.0289035). However, there is not a significant difference between the mean writing scores of students with SES levels 1 and 2 (p adj = 0.7096950).

We also conducted a Tukey pairwise comparision based on the bootstrap data.
```{r echo=FALSE, message=FALSE}
# conduct Tukey with the bootstrap data
# prepare data
df <- data.frame(ses = c(rep("low", length(the_sample1)), rep("medium", length(the_sample2)), rep("high", length(the_sample3))),
                 score = c(the_sample1, the_sample2, the_sample3))

# ANOVA
model2 <- aov(df$score ~ df$ses, data = df)
print(summary(model2))

# Tukey
tukey_data2 <- TukeyHSD(model2)
print(tukey_data2)
```
In this test, we made ses1 = "low", ses2 = "medium", ses3 = "high". The ANOVA table shows that there is a statistically significant difference in the mean scores of the groups defined by the variable ses at the alpha level of 0.05. The F-value is 4.004 and the p-value is 0.0197.

The Tukey multiple comparisons of means table shows the pairwise comparisons between the three groups (SES levels low, medium, and high). Based on the table, we can conclude that there is a statistically significant difference between the mean scores of students with SES levels low and high (p adj = 0.0151801), but not between students with SES levels medium and high (p adj = 0.1754145) or medium and low (p adj = 0.3350980).

Both Tukey pairwise comparision tests concluded statistically significant difference between the writing mean scores of students with ses level1 and level3. The Tukey test based on the original sample also concluded statistically significant difference between ses level2(medium) and level3(high).



