---
title: "Bootstrap Omnibus Test for Project G"
author: "Linda Wei"
output: html_notebook
---

#### Bootstrap Omnibus Test inference 

```{r echo=FALSE, message=FALSE}
# load the data
library(here)
school_data <- read.csv(here("data", "hsb2.csv"))

# use the bootstrap function with 3 factor samples to find the five number summary and CI

  ses1 <- subset(school_data, ses == 1)$write
  ses2 <- subset(school_data, ses == 2)$write
  ses3 <- subset(school_data, ses == 3)$write
  
  # bootstrap function
  mean_vector <- NULL
  n1 <- length(ses1)
  n2 <- length(ses2)
  n3 <- length(ses3)
  
  for (i in 1:10000) {
    the_sample1 <- sample(ses1, n1, replace = TRUE)
    mean_sample1 <- mean(the_sample1)
    the_sample2 <- sample(ses2, n2, replace = TRUE)
    mean_sample2 <- mean(the_sample2)
    the_sample3 <- sample(ses3, n3, replace = TRUE)
    mean_sample3 <- mean(the_sample3)
    the_sample <- c(the_sample1, the_sample2, the_sample3)
    mean_vector <- c(mean_vector, mean(the_sample))
  }
  
  mean_vector <- sort(mean_vector)
  boot_int <- c(mean_vector[251], mean_vector[9750])
  
# print the distributions
  fivenum_data <- fivenum(school_data$write)
  fivenum_boot <- fivenum(mean_vector)
  boot_int_data <-c(boot_int[1], NA, NA, NA, boot_int[2])

table_data <- matrix(c(fivenum_summary, fivenum_boot, boot_int_data), nrow = 3, byrow = TRUE,
                     dimnames = list(c("Five-number summary", "Five-number boot", "Bootstrap confidence interval"),
                                     c("Minimum", "1st Q", "Median", "3rd Q", "Maximum")))

print(table_data)
```

Based on the bootstrap result, we are 95% confident that the true population parameter on writing test score lies between `r boot_int[1]` and `r boot_int[2]`. The null hypothesis is there is no mean difference among the SES groups on writing test scores. The confidence interval does not contain a specific value (such as zero), we can conclude that the probability of observing a value as extreme as or more extreme than that value under the null hypothesis is less than the level of significance (e.g., 0.05). This provides evidence against the null hypothesis and suggests that the observed difference in means (or other population parameter) is statistically significant.

I aslo use the bootstrap function to calculate the F statistics and p-value to make inference
```{r echo=FALSE, message=FALSE}
# another way to make bootstrap inference based on F statistics
# compute the observed means and the omnibus F-statistic
obs_means <- tapply(school_data$write, school_data$ses, mean)
ss_between <- sum((obs_means - mean(school_data$write))^2)
df_between <- 2
ss_within <- sum((school_data$write - obs_means[school_data$ses])^2)
df_within <- 197
F_statistic <- (ss_between / df_between) / (ss_within / df_within)

print(F_statistic)

# compute the bootstrap means and F-statistic

boot_F_stats <- numeric(10000)
for(i in 1:10000) { 

boot_means <- tapply(the_sample, school_data$ses, mean)
boot_ss_between <- sum((boot_means - mean(the_sample))^2)
boot_ss_within <- sum((the_sample - boot_means[school_data$ses])^2)
df_between <- length(unique(school_data$ses)) - 1
df_within <- length(the_sample) - length(unique(school_data$ses))
boot_F_stats[i] <- (boot_ss_between / df_between) / (boot_ss_within / df_within)
}  

# Calculate the p-value

p_value <- sum(mean(boot_F_stats) >= F_statistic) / length(boot_F_stats)

print(p_value)

```

Based on the above bootstrap test on F statistics and p-value, the p-value is less than .05 which indicated that at 95% confidence level there are mean differences among different SES groups on the writing test scores. 

#### Tukey pairwise confidence interval

Based on the bootstrap test result, the overall CI does not contain zero, it indicates that there is a statistically significant difference between the means of the three ses groups. However, it does not provide information on which specific pairs of means differ significantly from each other.Therefore, a post-hoc test Tukey pairwise comparison (Tukey HSD test) is conducted to determine which specific pairs of means differ significantly from each other, while controlling for the overall type I error rate.

First, a Tukey pairwise test is conducted based on the hsb2.csv data.
```{r echo=FALSE, message=FALSE}
# prepare data for ANOVA
write_data <- school_data[, c(4, 8)]

write_data$ses <- factor(write_data$ses,
                          levels = c(1, 2, 3))

# ANOVA
model <- aov(write_data$write ~ write_data$ses, data = write_data)
print(summary.aov(model))

# TukeyHSD
tukey_data <- TukeyHSD(model)

print(tukey_data)
```
The ANOVA table shows that there is a statistically significant difference in the mean writing scores of the groups defined by the variable ses (which stands for socioeconomic status) at the alpha level of 0.05. The F-value is 4.97 and the p-value is 0.00784.

The Tukey multiple comparisons of means table shows the results of pairwise comparisons between the three groups (SES levels 1, 2, and 3). Based on the table, we can conclude that there is a statistically significant difference between the mean writing scores of students with SES levels 1 and 3 (p adj = 0.0114079) and between students with SES levels 2 and 3 (p adj = 0.0289035). However, there is not a significant difference between the mean writing scores of students with SES levels 1 and 2 (p adj = 0.7096950).

We also conducted a Tukey pairwise comparision based on the bootstrap data.
```{r echo=FALSE, message=FALSE}
# conduct Tukey with the bootstrap data
# prepare data
df <- data.frame(ses = c(rep("low", length(the_sample1)), rep("medium", length(the_sample2)), rep("high", length(the_sample3))),
                 score = c(the_sample1, the_sample2, the_sample3))

# ANOVA
model2 <- aov(df$score ~ df$ses, data = df)
print(summary(model2))

# Tukey
tukey_data2 <- TukeyHSD(model2)
print(tukey_data2)
```
In this test, we made ses1 = "low", ses2 = "medium", ses3 = "high". The ANOVA table shows that there is a statistically significant difference in the mean scores of the groups defined by the variable ses at the alpha level of 0.05. The F-value is 4.004 and the p-value is 0.0197.

The Tukey multiple comparisons of means table shows the pairwise comparisons between the three groups (SES levels low, medium, and high). Based on the table, we can conclude that there is a statistically significant difference between the mean scores of students with SES levels low and high (p adj = 0.0151801), but not between students with SES levels medium and high (p adj = 0.1754145) or medium and low (p adj = 0.3350980).

Both Tukey pairwise comparision tests concluded statistically significant difference between the writing mean scores of students with ses level1 and level3. The Tukey test based on the original sample also concluded statistically significant difference between ses level2(medium) and level3(high).



